---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Andrey Samarskiy (12252674)"
date: "`r Sys.Date()`"
editor_options: 
  markdown: 
    wrap: 72
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.

-   Include all your team-member's names and student numbers
    (Martrikelnummern) in the `authors` field.

-   Please use the exercise template document to work on and submit your
    results.

-   Use a level 2 heading for each new exercise and answer each subtask
    next to it's bullet point or use a new level 3 heading if you want.

-   Always render the R code for your solutions and make sure to include
    the resulting data in your rendered document.

    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).

-   Always submit both the rendered document(s) as well as your source
    Rmarkdown document. Submit the files separately on moodle, **not**
    as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)

    *Strengths*: 

    1\) Works as a back-up tool to ensure that you can always regain
    access to a lost file or correct unwanted changes.

    2\) Facilitates collaboration across devices (no more sending
    scripts back and forth). Also allows a straightforward
    identification of the changes your colleagues made.

    *Weaknesses*:

    1\) A file's history requires additional storage (albeit very
    little).

    2\) Doesn't work well with binary files (e.g., PDFs) and large data
    sets.

e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.

    ```{r}
    #library(rmarkdown)
    #render("ExerciseSheet_3_TEMPLATE.Rmd")
    ```

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:

    git\@github.com:asamarskiy/exeRcise-sheet-3.git

    (alternative: <https://github.com/asamarskiy/exeRcise-sheet-3.git>)

b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.

c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

    Text for push test

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.

    ```{r}
    library(tidyverse)

    nam21 <- read_csv("data/vornamen_2021.csv")
    nam22 <- read_csv("data/open_data_portal_2022.csv")

    typeof(nam21$Anzahl)
    typeof(nam22$Anzahl)
    ```

    The data contain some entries of the following form:

    ```{r}
    nam22$Anzahl[495]
    ```

    Since names that appear four times or less contain non-numeric
    characters, the entire variable "Anzahl" receives the type
    character.

    For this cases, I assume (arbitrarily) that the average count within
    the "less than 4" subset is 2. This motivates the following
    transformations.

    ```{r}
    nam21 <- nam21 %>% mutate(Anzahl = str_replace(Anzahl, "4 oder weniger", "2")) %>% 
            mutate(Anzahl = as.numeric(Anzahl))

    nam22 <- nam22 %>% mutate(Anzahl = str_replace(Anzahl, "4 oder weniger", "2")) %>% 
            mutate(Anzahl = as.numeric(Anzahl))
    ```

    *Discussion*. The solution successfully transforms the variable
    types.

    ```{r}
    typeof(nam21$Anzahl)
    typeof(nam22$Anzahl)
    ```

    But I essentially guess the count of rare names (≤ 4) by setting it
    equal to 2 in all cases. This arbitrary assumption is problematic
    since rare names make up the majority of given names (more than 80%
    in 2021):

    ```{r}
    placeh <- nam21 %>% mutate(rare = ifelse(Anzahl == 2, 1, 0))
    sum(placeh$rare)/length(placeh$rare)
    ```

c)  Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?

    ```{r}
    c(sum(nam21$Anzahl), sum(nam22$Anzahl))
    ```

    2021 had more births in Munich take place.

d)  Add a new column `year` to both datasets which holds the correct
    year for each.

    ```{r}
    nam21 <- nam21 %>% mutate(year = 2021)
    nam22 <- nam22 %>% mutate(year = 2022)
    ```

e)  Combine both datasets into one using `bind_rows()`.

    ```{r}
    names <- bind_rows(nam21, nam22)
    ```

f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.

    ```{r}
    names <- names %>% group_by(Vorname) %>% mutate(count_both = sum(Anzahl))

    tabledat <- names %>% arrange(desc(count_both)) %>%
              distinct(Vorname, .keep_all = TRUE)
      
    knitr::kable(head(tabledat[, c("Vorname", "count_both")], 10), "pipe", caption = "Top 10 most popular baby names in Munich (2021-22)")
              
    ```

## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

a)  The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

###Create a sample data frame

tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) )

###Calculate column-wise means

column_means \<- lapply(tax_data\[, -1\], 2, mean)

column_means

*Prompt*: I'll send you an R code chunk. There is a mistake because the
wrong apply function has been used. Correct the mistake and explain it.
Here's where it gets tricky: You can only replace the lapply function.
The replacement must come from the apply() function collection. Further,
you must keep its arguments (tax_data\[, -1\], 2, mean) unchanged. DO
NOT CHANGE THE ARGUMENTS UNDER ANY CIRCUMSTANCES. The code is correct
except for the function being used.

*No matter what I write, ChatGPT always fiddles with the arguments when
suggesting a different function. The above prompt worked only with this
follow-up:*

*Prompt*: Have you tried replacing lapply with apply?

*Response*: Apologies for the oversight. To replace `lapply()` with a
function from the `apply()` family, we can use `apply()` in combination
with the `MARGIN` argument set to 2 to calculate column-wise means.
Here's the corrected code using `apply()`:

```{r}
# Create a sample data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate column-wise means using apply()
column_means <- apply(tax_data[, -1], 2, mean)

column_means
```

In this code, `apply()` is used with `tax_data[, -1]` as the first
argument to specify the data frame from which the column-wise means
should be calculated. The `2` argument indicates that the calculation
should be performed on each column. Finally, `mean`
is used as the third argument to specify the function to apply to each
column. The result will be a vector of column means stored in the
`column_means` variable.

*Prompt:* So, what was the mistake in the original code?

*Response*: The mistake in the original code was the incorrect usage of
the `lapply()` function. The second argument of `lapply()` is expected
to be a function, but the value `2` was provided instead. The correct
usage of `lapply()` should be:

lapply(X, FUN, ...)

Where `X` is the input data structure and `FUN` is the function to be
applied to each element of `X`. In the original code, the argument `2`
was used instead of a function, resulting in an error.

*Prompt:* How would you calculate the rowwise means?

*Response*: suggested `rowMeans()`

*Prompt:* Can you use an alternative solution?

*Response*: Certainly! An alternative solution to calculate the row-wise
means is by using the `apply()` function with `MARGIN = 1`. Here's the
modified code using `apply()`:

```{r}
# Create a sample data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate row-wise means using apply()
row_means <- apply(tax_data[, -1], 1, mean)

row_means
```

In this alternative solution, `apply()` is used with `tax_data[, -1]` as
the first argument to specify the data frame from which the row-wise
means should be calculated. The `1` argument indicates that the
calculation should be performed on each row. Finally, `mean`
is used as the third argument to specify the function to apply to each
row. The result will be a vector of row means stored in the `row_means`
variable.

b)  Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.

*Prompt*: Can you explain what the rapply() function does in R? Please
illustrate with an example.

*Response*: In a nutshell, the `rapply()` function in R is used to apply
a given function to every element of a nested structure, such as a list.
It recursively goes through each element, including nested elements, and
performs the specified function on them.

*Example (Comments mine)*:

```{r}
# ChatGPT created a list with two numbers and another list, also containing two numbers.
numbers <- list(
  1,
  2,
  nested_numbers = list(
    3,
    4
  )
)

# We create the double function (*2 multiplication)
double <- function(x) {
  x * 2
}

#rapply permits us to double each number in the list, indepently of whether it is nested or not
result <- rapply(numbers, double)

print(result)
```

## Final Note

Make sure to push all your commits and changes to GitHub before
submittining the exercise sheet.
